---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RecurrenceEnd

<!-- badges: start -->
<!-- badges: end -->

The goal of RecurrenceEnd is to implement several non-parametric methods  for estimating the distribution of the unobservable terminating event of a recurrentevent process.

## Installation

You can install the development version of RecurrenceEnd from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("anikoszabo/RecurrenceEnd")
```

## Example

```{r example}
library(RecurrenceEnd)
```

The `SimulatedData` data set contains simulated recurrent events for 97 subjects. The recurrent events continue until an unobserved end-time, when they stop. The goal is to estimate the distribution of the unobserved end-time across the subjects.

The data has one row for each occurrence of the recurrent event within a subject, and a last row for the follow-up time during which no additional events have occurred. The following table shows data from 3 subjects with varying numbers of recurrent events. The 'time' variable records the event/last follow-up time, the 'indicator' distinguishes between events (1) and last follow-up (0), while Z.1 and Z.2 are covariates that affect the rate of the recurrent event process.

```{r showdata}
SimulatedData[SimulatedData$patient.id %in% c(5,6,7),]
```


We can visualize the data to get a better understanding of its structure. In the plot, each gray line represents the follow-up of a patients, with the blue points showing the occurrences of the recurrent event in that patient. We can see that for some subjects the recurrent events seem to keep going until the last follow-up time (and likely beyond it as well), while for others there is a long stretch of follow-up time without any recurrent events, indicating that the unobserved ending event has probably happened.


```{r plot}
library(reReg) # for plotting recurrent event data
plotEvents(Recur(time=time, id=patient.id, event=indicator) ~ 1,
           data=SimulatedData, 
           control=list(cex=1, width=0.2, recurrent.color="navy")) 
```

This package implements four methods for estimating the distribution of the unobserved
end-time. The NPMLE method is recommended, the others are ad-hoc methods implemented for comparison.

```{r FitModel}
res_np <- estimate_end(Recur(time=time, id=patient.id, event=indicator) ~ Z.1+Z.2, 
                       data = SimulatedData, bootCI = TRUE, bootB = 100)

plot(res_np, col="blue", conf.int = TRUE, conf.lty=3, conf.col="blue")
```

The values at specific time-points can be obtained using the `predict` function, while quantiles can be extracted using the `quantile` and `median` functions.

```{r Estimates}
# probability of ongoing recurrent events at time = 1
predict(res_np, times=1, conf.int = TRUE)

# median ending time
median(res_np, conf.int=TRUE)
```

